---
title: '2019 Pittsburgh Steelers: What Are The Keys to victory in 2020?'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
Steelers_Data <- read_excel("C:/Users/student/OneDrive - South Dakota State University - SDSU/Junior Classes/STAT 442/Steelers Data.xlsx")
steelers <- Steelers_Data
summary(steelers)
```

Some random stats:
Record: 8-8
Lost first 3 games and last 3 games of the season. They were 8-2 in between. 
Combined record of teams they played when they faced them: 56-63-1 (.475)
Combined record of teams they played at the end of the season: 128-127-1 (.502)
Total Point For: 289
Total Points Against: 303
Total Yards For: 4428
Total Yards Against: 4866
Turnover Ratio: +8 (38-30)
Led the league in sacks (54)



*** Making data sets separating games they won and games they lost ***

```{r}
library(dplyr)
win <- steelers %>% filter(Win_Loss == "Win")
win

loss <- steelers %>% filter(Win_Loss == "Loss")
loss
```
Stats from looking at means from the summaries of wins and losses setss:
Points scored: 22.5 in wins, 13.6 in losses
Points Against: 13.8 in wins, 24.1 in losses
Rush yards For: 113 in wins, 68 in losses     ****
Rush yards Against: 82 in wins, 137 in losses
3rd down Percentage For: 40% in wins, 28% in losses
3rd down Percentage Against: 31% in wins, 41% in losses     ****    








```{r}
steelers %>%
  summarise_at(vars(Opp_Overall_Record),
               list(name=mean))
win %>%
  summarise_at(vars(Opp_Overall_Record),
               list(name=mean))
loss %>%
  summarise_at(vars(Opp_Overall_Record),
               list(name=mean))
```

Bar plots to see what variables effected winning/losing games

```{r}
library(ggplot2)
library(plotly)
p <- ggplot(steelers, aes(x = reorder(Opponent, Rush_Yards_For), y = Rush_Yards_For)) +
               geom_col(aes(fill=Win_Loss)) + coord_flip()
ggplotly(p)

p <- ggplot(steelers, aes(x = reorder(Opponent, Rush_Yards_Against), y = Rush_Yards_Against)) +
                   geom_col(aes(fill=Win_Loss)) + coord_flip()
ggplotly(p)

p <- ggplot(steelers, aes(x = reorder(Opponent, Sacks), y = Sacks)) +
                     geom_col(aes(fill=Win_Loss)) + coord_flip()
ggplotly(p)

p <- ggplot(steelers, aes(x = Third_Down_For_Perc, y = Third_Down_Against_Perc)) +
     geom_point(aes(color=Win_Loss)) + coord_flip()
ggplotly(p)

p <- ggplot(steelers, aes(x = reorder(Opponent, Off_Turnovers), y = Off_Turnovers)) +
  geom_segment( aes(x=Opponent, xend=Opponent, y=0, yend=Off_Turnovers, color = Win_Loss)) +
  geom_point( aes(color=Win_Loss), size=4, alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplotly(p)

p <- ggplot(steelers, aes(x = reorder(Opponent, Def_Turnovers), y =Def_Turnovers)) +
  geom_segment( aes(x=Opponent, xend=Opponent, y=0, yend=Def_Turnovers, color = Win_Loss)) +
  geom_point( aes(color=Win_Loss), size=4, alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )
ggplotly(p)
```


Making a correlation plot with all numeric variables
```{r}
library(ellipse)
library(RColorBrewer)
correlation <- Steelers_Data
correlation$Week = NULL
correlation$Opp_Record = NULL
correlation$Opponent = NULL
correlation$Win_Loss = NULL
correlation$Third_Down_Against = NULL
correlation$Third_Down_For = NULL
correlation
data <- cor(correlation)
my_colors <- brewer.pal(5, "Spectral")
my_colors <- colorRampPalette(my_colors)(100)
ord <- order(data[1, ])
data_ord <- data[ord, ord]
plotcorr(data_ord , col=my_colors[data_ord*50+50] , mar=c(0,0,0,0))
```

```{r}
p <- ggplot(steelers, aes(x = Sacks, y = Points_For)) +
                                   geom_point(aes(color=Win_Loss)) + coord_flip()
ggplotly(p)

p <- ggplot(steelers, aes(x = Def_Turnovers, y = Points_For)) +
                                   geom_point(aes(color=Win_Loss)) + coord_flip()
ggplotly(p)

cor(steelers$Sacks, steelers$Points_For)
cor(steelers$Def_Turnovers, steelers$Points_For)
```

*** Keys to Victory for the 2020 Pittsburgh Steelers ***

1. Rush Yards > 100 and allow < 100 rush yards
        5-0 when reaching 100 yards, 2-7 when giving up 100
2. Sacks >= 4
        7-2 when recording 4 sacks or more
3. Win the turnover battle
        8-3 when winning turnover battle
4. 3rd Down % > 33% and allow < 33% 3rd Down Percentage
        5-0 when both these things occur
      
Passing yards did not have a big effect on wins/losses




